Notes on Perturbational Quantities

### 2008-07-30 docs/pert_notes/paper.tex
this doc is in good shape now

-w_0\frac{\partial\rho_0}{\partial r} - \rho_0\frac{\partial w_0}{\partial r}.

@@

Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange, noting that 

@@

\subsubsection{Predicting $\rho_0$ at edges}
There are two ways to predict $\rho_0$ at edges.\\ \\
We can call make\_edge\_state\_1d using the term proportional to $\rho_0$ in 

@@

This term is used to advect $\rho_0$ in advect\_base and to compute $\eta_\rho$ that will be 
used in correct\_base.\\ \\
We can also define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We first 
compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which evolves (\ref{rho0 equation}) 
through $\Delta t$ in time.  The $(2)$ in the superscript indicates that we have not called 
correct\_base yet (correct\_base takes $\rho_0^{(2)}$ as an input and computes $\rho^{n+1}$ 
as the output).  This term is used to construct $\rho$ at edges from $\rho'$ at edges, and 
this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  We use $\rho^{(2)}$ rather 
than $\rho^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be consistent with the fact 
that we do not include an $\eta_\rho$ contribution to compute $\rho'$ at edges.

@@

For the non-radial edges, we can directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 

@@

\subsubsection{Advancing $\rho X_k$}

@@

-\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}}+\rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.

@@

The full temperature equation is:

@@

These are working notes for how we are going to handle the energy evolution.  Due to the 
operator splitting, the call to react\_state has already been made.  Hence, the goal is to 
compute an edge state energy starting from $(\rho h)^{(1)}$ using an energy evolution 
equation that does not include terms already accounted for in react\_state

@@

We can arbitrarily define a base state enthalpy evolution equation:

@@

&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 

@@

Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) gives the perturbational 
enthalpy equation:

@@

{\bf What the code is currently doing:} Instead of defining 
$(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries through the complete 
evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ react\_base 

@@

to sync up $(\rho h)_0$ after the thermal diffusion setp.

@@

@@

We first compute $(\rho h)_0^{(2)}$ from $(\rho h)_0^{(1)}$ using advect\_base, which evolves 
(\ref{rhoh equation}) through $\Delta t$ in time.  Then we follow an analogous procedure 
used in Section \ref{Computing rho at edges} to obtain $\rho h$ at edges.

@@

We would like to update the non-forcing part of the enthalpy equation using

@@

\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = -\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}}+(\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\}.

@@

{\bf What the code is currently doing:}

@@

\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt + w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\right].

>>>>

-w_0\frac{\partial\rho_0}{\partial r} 
\underbrace{-\rho_0\frac{\partial w_0}{\partial r}}_{"\rho_0 ~ \text{force}"}.

@@

Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange terms, noting that 

@@

\subsubsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
There are two ways to predict $\rho_0$ at edges.
\begin{enumerate}
\item We call make\_edge\_state\_1d using the underbraced term in 

@@

This term is used to advect $\rho_0$ in advect\_base and to compute $\eta_\rho$, which will 
be used in correct\_base.
\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We 
compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which advances equation 
(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the superscript indicates 
that we have not called correct\_base yet, which computes  $\rho^{n+1}$ from $\rho_0^{(2)}$. 
We use $\rho^{(2)}$ rather than $\rho^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be 
consistent with the fact that we do not include an $\eta_\rho$ contribution to compute 
$\rho'$ at edges.  $\rho_0^{n+\half,{\rm avg}}$ is used to construct $\rho$ at edges from 
$\rho'$ at edges, and this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  
\end{enumerate}

@@

For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 

@@

\subsubsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}

@@

-\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}} 
+ \rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.

@@

The full temperature equation is

@@

Due to operator splitting, the call to react\_state has already been made.  Hence, the goal 
is to compute an edge state enthalpy starting from $(\rho h)^{(1)}$ using an enthalpy
equation that does not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where were 
already accounted for in react\_state

@@

We arbitrarily define a base state enthalpy evolution equation
\MarginPar{Does $\psi$ need to be included in the force?}

@@

&=& -w_0\frac{\partial(\rho h)_0}{\partial r} 
- \underbrace{(\rho h)_0\frac{\partial w_0}{\partial r}}_{"(\rho h)_0 ~ \text{force}"}

@@

Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and rearranging terms gives 
the perturbational enthalpy equation

@@

{\bf What the code is currently doing:}\MarginPar{Do we need to change this?}
 Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ react\_base 

@@

to sync up $(\rho h)_0$ after the thermal diffusion step.

@@

We use an analogous procedure described in Section \ref{Predicting rho0 at edges} to 
obtain $(\rho h)_0^{n+\half,\rm{pred}}$ and $(\rho h)_0^{n+\half,\rm{avg}}$.  Note that 
$(\rho h)_0^{n+\half,{\rm avg}} = [(\rho h)_0^{(1)} + (\rho h)_0^{(2)}]/2$.

@@

We use an analogous procedure described in Section \ref{Computing rho at edges} to compute
$\rho h$ at edges.

@@

We would like to update the non-forcing part of the enthalpy equation analogously to
Section \ref{Advancing rhoX_k}

@@

\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = 
-\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}} 
+ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\}.

@@

{\bf What the code is currently doing:}\MarginPar{Do we need to change this?}

@@

\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt 
+ w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\right].
